Group 9 Presentation

### 2021-12-06 Bagging Model.Rmd
Updating models

library(readr)
heart <- read_csv("heart.csv")
View(heart)

#60/40 split
heart_df <- sample(x=nrow(heart), size=.40*nrow(heart))
train <- heart[heart_df,]
test <-  heart[-heart_df,]

@@

```{r}

bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500) 
predict_bag <- predict(bagging, newdata = test)

mean((predict_bag - test$HeartDisease)^2)

@@

Now trying to visulize the bagging

```{r}

VI <- data.frame(var=names(heart[,-1]), imp=varImp(bagging))

#sort variable importance descending
VI_plot <- VI[order(VI$Overall, decreasing=TRUE),]

>>>>

library(readr)

#60/40 split
train <-train
test <- test

@@

```{r}
bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
                        importance = TRUE,
                        proximity = TRUE)

#Confusion Matrix
predict_bag <- predict(bagging, newdata = test)
predicted_bag <- ifelse(predict_bag>0.5,1,0)
confm_bag<-table(predict_bag,test$HeartDisease)
confm_bag

#MSE
bagging_mse <- mean((predict_bag - test$HeartDisease)^2)

#Over all correct:
overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)

kable(confm_bag) #confusion matrix
overall_bag #over all correct

#Visulizaton of bagging
VI <- data.frame(var=names(test[,-1]), imp=varImp(bagging))
VI
#sort variable importance descending
VI_plot <- VI[order(VI$Overall, decreasing=TRUE),]

@@

### 2021-12-06 KNN Model.Rmd
Updating models

---
title: "Group 9 Presentation"
author: "Henrique, Oriana, Jorie"
date: "12/4/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(repos='http://cran.rstudio.org')
have.packages <- installed.packages()
cran.packages <- c('devtools','plotrix','randomForest','tree')
to.install <- setdiff(cran.packages, have.packages[,1])
if(length(to.install)>0) install.packages(to.install)

library(devtools)
if(!('reprtree' %in% installed.packages())){
  install_github('araastat/reprtree')
}
for(p in c(cran.packages, 'reprtree')) eval(substitute(library(pkg), list(pkg=p)))
```

```{r, include=FALSE}

#reproducibility
set.seed(445)

library(ISLR)
library(leaps)
library(tidyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls) 
library(tree)
library(randomForest)
library(knitr)
library(class)
library(fastDummies)
library(reprtree)

train <- read.csv("train.csv")
test <- read.csv("test.csv")

train$ChestPainType <- as.factor(train$ChestPainType)
train$Sex <- as.factor(train$Sex)
train$FastingBS <- as.factor(train$FastingBS)
train$RestingECG <- as.factor(train$RestingECG)
train$ExerciseAngina <- as.factor(train$ExerciseAngina)
train$ST_Slope <- as.factor(train$ST_Slope)
train$HeartDisease <- as.factor(train$HeartDisease)

test$ChestPainType <- as.factor(test$ChestPainType)
test$Sex <- as.factor(test$Sex)
test$FastingBS <- as.factor(test$FastingBS)
test$RestingECG <- as.factor(test$RestingECG)
test$ExerciseAngina <- as.factor(test$ExerciseAngina)
test$ST_Slope <- as.factor(test$ST_Slope)
test$HeartDisease <- as.factor(test$HeartDisease)

## Group 9

Heart Disease
Oriana Meldrum, Henrique Rio, Jorie Alvis

## LDA

Jorie's LDA

## RF

Henrique's RF

```{r}
rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)-1), importance = TRUE, type= "classification")

true <- test$HeartDisease
rf <- predict(rf_fit,test,type="class")

predicted2 <- ifelse(rf>0.5,1,0)

confm2<-table(predicted=rf,True=true)
kable(confm2)

paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))

rf_fit
```

## Representative Tree
-Following the paper by  Banerjee, et al.(2011) and the package reprtree we can get a representative tree from the random forest.
-This tree is calculated using a measure of similarity in which 2 trees are similar if they have the same prediction for all observations.
-Using this idea on the random forest we get the following tree:

```{r}
reptree <- ReprTree(rf_fit, train, metric='d2')
plot(reptree, index=1)

## SVM

Jorie's SVM

## (the new thing)

Henrique's new thing insted of PCA

## Logistic Regression

Henrique's Logisic Regression

```{r}
model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
#summary(model_glm)

predicted2 <- predict(model_glm, newdata = test, "response")
predicted2 <- ifelse(predicted2>0.5,1,0)
confm2<-table(predicted=predicted2,True=test$HeartDisease)
kable(confm2)

paste("The overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))

 #MSE

## Boosting

Jories's Boosting

## LASSO

Oriana's Lasso

@@

train_matrix <- model.matrix(HeartDisease ~ ., data = train)
test_matrix <- model.matrix(HeartDisease ~ ., data = test)
grid <- 10^seq(10, -2, length=100)

lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
#plot(lasso)

cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
#plot(cv_lasso)
best_lambda_lasso <- cv_lasso$lambda.min

lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)

num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]

# MSE values to use:
best_lambda_lasso
lasso_MSE
num_zeros

#Confusion Matrix
#predicted_lasso <- predict(lasso, newdata = test, "response")
#predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
#confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
#confm_lasso

#Over all correct:
#overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)

#kable(confm_lasso) #confusion matrix
#overall_lasso #over all correct
```

## KNN

@@

#trainknn <- temp_train[-temp_train$HeartDisease]
#testknn <- test[-test$HeartDisease]

num_k <- seq(1:100)

my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)

@@

overall_knn_1 <- (confm_knn_1[1,1]+confm_knn_1[2,2])/sum(confm_knn_1)

@@

 #confusion matrix
kable(confm_knn_1)
mean(knn_pred_1 !=test$HeartDisease)
```

## Bagging
```{r}

bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
                        importance = TRUE,
                        proximity = TRUE) 
predict_bag <- predict(bagging, newdata = test)

#MSE
bagging_mse <- mean((predict_bag - test$HeartDisease)^2)

#Confusion Matrix
predicted_bag <- predict(bagging, newdata = test)
predicted_bag <- ifelse(predicted_bag>0.5,1,0)
confm_bag<-table(predicted=predicted_bag,True=test$HeartDisease)
confm_bag

#Over all correct:
overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)

kable(confm_bag) #confusion matrix
overall_bag #over all correct

@@

## Comparision of all method's MSE

Chart with all MSE

```{r}
library(knitr)
df1 = data.frame(Method = c('LDA', 'RF', 'SVM', 'new THING', 'Logistic Regression', 'Boosting', 'LASSO', 'PLS', 'Bagging'),  MSE = c(0,0,0,0,0.1498638,0,0,0.1054 ,0.1165492))

kable(df1)

## Conclusion

the METHOD worked best, with an MSE of SOMETHING

```{r pressure}
plot(pressure)
```

>>>>

---
title: "KNN Model"
author: "Oriana Meldurm"
date: "12/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

@@

library(ISLR)
library(class)
library(tidyr)

install.packages("fastDummies")
library(fastDummies)

@@

trainknn <- temp_train[-temp_train$HeartDisease]
testknn <- test[-test$HeartDisease]

trainknn
testknn

?knn.cv
num_k <- seq(1:100)

my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)
my_cv

@@

overall_knn_1 <- (confm_knn[1,1]+confm_knn[2,2])/sum(confm_knn)

@@

 #confusion matrix
kable(confm_knn)
mean(knn_pred !=test$HeartDisease)

@@

library(caret)

#5 fold
trControl <- trainControl(method = "cv", number = 5)

fit <- train(as.factor(HeartDisease) ~.,
             method = "knn", 
             tuneGrid = expand.grid(k = 1:100), 
             trControl = trControl, 
             metric = "Accuracy", 
             data = temp_train)

fit
```

### 2021-12-06 Lasso Model.Rmd
Updating models

lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12)
plot(lasso)

#Here it breaks
#cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
#plot(cv_ridge)
#best_lambda_lasso <- cv_lasso$lambda.min

#lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
#lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)

#num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]

#values to use:
#best_lambda_lasso
#lasso_MSE
#num_zeros

>>>>

lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
#plot(lasso)

cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
#plot(cv_lasso)
best_lambda_lasso <- cv_lasso$lambda.min

lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)

num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]

# MSE values to use:
best_lambda_lasso
lasso_MSE
num_zeros

#Confusion Matrix
predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
confm_lasso

#Over all correct:
overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)

kable(confm_lasso) #confusion matrix
overall_lasso #over all correct

### 2021-12-06 PLS Model.Rmd
Updating models

>>>>

## PLS

pls_fit <- plsr(HeartDisease ~ ., data = train, type= "class", scale = TRUE, validation = "CV")
#validationplot(pls_fit, val.type = "MSEP")

pls_prediction <- predict(pls_fit, test,  ncomp = 7)
#pls_MSE <- mean((test[, "HeartDisease"] - data.frame(pls_prediction))^2)

#Confusion Matrix prep
predicted_pls <- predict(pls_fit, newdata = test)
predicted_pls <- ifelse(predicted_pls>0.5,1,0)
#confm_pls<-table(predicted=predicted_pls,True=test$HeartDisease)

#Over all correct:

#overall_pls <- (confm_pls[1,1]+confm_pls[2,2])/sum(confm_pls)

#values to use:
#mse
#m_pls

#kable(confm_pls) #confusion matrix
#overall_pls #over all correct

### 2021-12-06 presentation.Rmd
Commit
Updating models

---
title: "Group 9 Presentation"
author: "Henrique, Oriana, Jorie"
date: "12/4/2021"
output: ioslides_presentation
---

@@

library(ISLR)
library(leaps)
library(tidyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls) 
library(tree)
library(randomForest)
library(knitr)
library(class)
library(fastDummies)
library(reprtree)

@@

@@

@@

cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
#plot(cv_lasso)
best_lambda_lasso <- cv_lasso$lambda.min

@@

#Confusion Matrix
#predicted_lasso <- predict(lasso, newdata = test, "response")
#predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
#confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
#confm_lasso

#Over all correct:
#overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)

#kable(confm_lasso) #confusion matrix
#overall_lasso #over all correct
```

@@

 #confusion matrix
kable(confm_knn_1)
mean(knn_pred_1 !=test$HeartDisease)

@@

bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
                        importance = TRUE,
                        proximity = TRUE) 
predict_bag <- predict(bagging, newdata = test)

@@

#Confusion Matrix
predicted_bag <- predict(bagging, newdata = test)
predicted_bag <- ifelse(predicted_bag>0.5,1,0)
confm_bag<-table(predicted=predicted_bag,True=test$HeartDisease)
confm_bag

@@

@@

```{r pressure}
plot(pressure)
```

>>>>

---
title: "Group 9 Presentation"
author: "Henrique, Oriana, Jorie"
date: "12/4/2021"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

@@

library(ISLR)
library(leaps)
library(tidyr)
library(dplyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls) 
library(tree)
library(randomForest)
library(knitr)
library(class)
library(fastDummies)
library(reprtree)

@@

## Introduction

In 2019 659,041 Americans died of heart disease. Four out of five heart disease deaths are due to heart attacks and strokes, and one-third of these deaths occur prematurely in people under 70 years of age. Heart failure is a common event caused by heart disease and this dataset we used contains 11 variables that can be used to predict a possible heart disease.

People with cardiovascular disease or who are at high cardiovascular risk (due to the presence of one or more risk factors) need early detection and management so machine learning techniques are of great clinical intrest for early detection.

(https://www.cdc.gov/nchs/fastats/leading-causes-of-death.htm, https://www.kaggle.com/fedesoriano/heart-failure-prediction)

@@

paste("The overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))

@@

cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
#plot(cv_lasso)
best_lambda_lasso <- cv_lasso$lambda.min

@@

#Confusion Matrix
predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
confm_lasso

#Over all correct:
overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)

kable(confm_lasso) #confusion matrix
overall_lasso #over all correct
```

@@

 #confusion matrix
kable(confm_knn_1)
mean(knn_pred_1 !=test$HeartDisease)
```

@@

bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
                        importance = TRUE,
                        proximity = TRUE)

#Confusion Matrix
predict_bag <- predict(bagging, newdata = test)
predicted_bag <- ifelse(predict_bag>0.5,1,0)
confm_bag<-table(predict_bag,test$HeartDisease)
confm_bag

@@

@@

#Visulizaton of bagging
VI <- data.frame(var=names(test[,-1]), imp=varImp(bagging))

#sort variable importance descending
VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]

#visualize variable importance with horizontal bar plot
barplot(VI_plot$imp.0,
        names.arg=rownames(VI_plot),
        horiz=TRUE,
        col='steelblue',
        xlab='Variable Importance')

@@
